Real-space computation of $E$/$B$-mode maps I: Formalism, Compact Kernels, and Polarized Filaments

### 2018-12-10 articles/jcap_referee_response/modified article jcap/article.tex
added final resubmitted version of the draft

The radial function ${}_\mm f(\beta)$ is oscillatory, regardless of the maximum $\ell_{\rm max}$. The envelope of the oscillation decays with angular separation at intermediate $\beta$ but increases as $\beta \rightarrow \pi$.  We find that the positive envelope of ${}_{\mm}f$ scales as $\beta^{-1.6}$ at intermediate $\beta$.  However, when we average over several oscillations with a moving window, we find that the average behavior at intermediate $\beta$ scales as $\beta^{-2}$, visible in \fig{fig:fbeta_avg}. \revisit{To take the average, we set the width of the moving window to $0.3\beta_0$, so that it narrows as the band limit increases, keeping the number of oscillations in the window roughly the same.  The finding of $\beta^{-2}$} agrees with the hypothesis for the {flat sky and continuum case}, for which e.g. Zaldarriaga (2001) \cite{Zaldarriaga2001a} argued that form of the radial function has to be ${}_\mm f(\beta>0)\simeq\beta^{-2}$ to ensure that the Fourier modes of Stokes $Q/U$ relate to those of $E/B$ merely by rotations (with no scale dependence). In addition, we find that ${}_{\md}f$ under a moving average shows a similar scaling of $\beta^{-2}$, as seen in \fig{fig:dfbeta_avg}.
\revisit{When the moving average is applied radial functions with progressively higher band limits, we see that a larger and larger fraction of angular domain matches the scaling of $\beta^{-2}$.}   Extrapolating this trend to the case of a very high band limit one expects this behavior to approach $\beta^{-2}$ across the domain $\beta\in[\epsilon,\pi-\epsilon]$ for $\epsilon \ll 1$. In essence the moving average can be understood as the averaging resulting from discrete sampling of a continuum field (one with an infinite band limit) where the width of the averaging window corresponds to the size of the pixel.

\revisit{In \fig{fig:fbeta_avg}, near the edges ($\beta \rightarrow 0,\pi$), the window function partly falls outside the domain, and this in part causes the the scaling to deviate from the power law $\beta^{-2}$.}  Note also that ${}_{\mm}f$ and ${}_{\md}f$ need to vanish at $\beta=0$ and therefore the sliding average in the vicinity of these points cannot have the same power law behavior.
%---------------------------------------------------------------------------------------------------------
%\include{generalizedeb}
\section{Generalized polarization operators and recovering standard power spectra}\label{sec:generalized_operators}  
\addtext{In this section we relate our new-found knowledge about the radial part of the kernel to the computation of the standard CMB polarization power spectra.} With this better understanding, we can write down generalized $E/B$-like fields that depend on a different radial function, even one that we specify to have compact support.
The spin symmetry constrains the the azimuthal part of the real space kernels to be of the form $\sim e^{\pm i2 \alpha}$.  The radial parts of the standard operators are determined by the sum over spherical harmonics and varies as a function of the band limit. It is here that we may potentially choose alternate forms for the radial functions to suit certain kind of analysis.

We can systematically generalize the real space operator by introducing the following harmonic space filter function:
%
\beq
\tilde{\mathcal{G}} = {\begin{bmatrix} g_{\ell}^E & 0  \\  0 & g_{\ell}^B \end{bmatrix}} \,,
\eeq
%
where the functions $g_{\ell}^E$ and $g_{\ell}^B$ represent the harmonic representation of the modified radial functions and can in the most general case be chosen to be different for $E$ and $B$ modes. To simplify discussions, we proceed by setting $g_{\ell}^E = g_{\ell}^B= g_{\ell}$. Given this harmonic function $g_{\ell}$, we can define the real space operator $\bar{O}'$ which translates Stokes $Q/U$ to $E/B$-like scalars (and the inverse operator $\bar{O}'^{-1}$) in the following manner,
% 
\begin{subequations} \label{eq:gen_qu2eb}
\beqry
{\bar O}' &=& {{}_0\mathcal{Y}} \, \tilde T^{-1} \tilde{\mathcal{G}} {{}_2\mathcal{Y}^{\ddag}} \, \bar T \,,\\
{\bar O}'^{-1}&=& \bar{T}^{-1} {{}_2\mathcal{Y}}\, \tilde{\mathcal{G}}^{-1} \tilde T {{}_0\mathcal{Y}^{\ddag}}.
\eeqry
\end{subequations}
%
The primed notation distinguishes these generalized operators from the default operators defined in \sec{sec:qu2eb} and \sec{sec:eb2qu}. We require both the forward and inverse operators to be well defined.   This constrains the choice of $\tilde{\mathcal{G}}$ to have a valid  inverse, which is important when recovering the standard CMB power spectra. The radial parts of this generalized operator and it's inverse are given by the following expressions,
%
\begin{subequations}\label{eq:generalized_radial_kernel}
\beqry 
G_{QU \rightarrow EB}(\beta) &=& G(\beta) = \sum _{\ell=2} ^{\ell_{\rm max}} g_{\ell}\frac{2 \ell+1}{4 \pi} \sqrt{\frac{(\ell-2)!}{(\ell + 2)!}} P_{\ell}^2(\cos{\beta}) \, \label{eq:mod_rad_forward} \,, \\
G_{EB \rightarrow QU}(\beta) &=& G^{-1}(\beta) = \sum _{\ell=2} ^{\ell_{\rm max}} g_{\ell}^{-1}\frac{2 \ell+1}{4 \pi} \sqrt{\frac{(\ell-2)!}{(\ell + 2)!}} P_{\ell}^2(\cos{\beta}) \,. \label{eq:mod_rad_inverse}
\eeqry
\end{subequations}
%
The default radial function is just a special case resulting from the choice $\tilde{\mathcal{G}}=\mathbb{1}$ ($g_{\ell}=1$), in which case $\tilde{\mathcal{G}}^{-1}=\tilde{\mathcal{G}}$ and therefore $G^{-1}(\beta) = G(\beta)={{}_{\mm}f}$.
%
\begin{figure}[!t] 
\centering
\subfigure[\label{fig:gl_gbeta}]{\includegraphics[width=0.48\columnwidth]{Gbeta_for_different_gl_lmax1536.pdf}}
\subfigure[\label{fig:glbl}]{\includegraphics[width=0.48\columnwidth]{gl_for_different_gbeta_lmax1536.pdf}}
\subfigure[\label{fig:gl_bbeta}]{\includegraphics[width=0.48\columnwidth]{Bbeta_for_gl_lmax1536.pdf}}
\caption{\textit{Top left:} The green line depicts the default radial kernel $f(\beta)$ defined in \eq{eq:qu2eb_gen_kernel}, multiplied by an apodized step function $\theta(3 \beta_0)$. The blue and orange lines depict the modified radial function resulting the beam harmonics $b_{\ell}$ corresponding to Gaussian beams with fwhm=15 \& 12 arc-minutes respectively. The red curve depicts an example modified radial function: $G(\beta)=\mathcal{A} \beta^{n_o} \exp{\left[ -\left( {(\beta-\beta_s)}{/\sqrt{2} \sigma} \right)^{n_1} \right]}$ with parameters set to the following values $[n_0=1;\, \beta_s=0 ;\, \sigma = 0.004 ;\, n_1=1.5]$.  \addtext{\st{The black dashed curve depicts the band limited reconstruction of the modified radial function.}} \textit{Top right:} The harmonic representation of the respective radial functions as indicated by the legend. The dashed curves of the corresponding color depict the inverse of the harmonic functions. \textit{Bottom:}  The normalized beam function $b(\beta)$ evaluated from interpreting the respective harmonic functions as those corresponding to an effective beam applied to $E/B$. }
\label{fig:example_gbeta}
\end{figure}
%

>>>>

The radial function ${}_\mm f(\beta)$ is oscillatory, regardless of the maximum $\ell_{\rm max}$. The envelope of the oscillation decays with angular separation at intermediate $\beta$ but increases as $\beta \rightarrow \pi$.  We find that the positive envelope of ${}_{\mm}f$ scales as $\beta^{-1.6}$ at intermediate $\beta$.  However, when we average over several oscillations with a moving window, we find that the average behavior at intermediate $\beta$ scales as $\beta^{-2}$, visible in \fig{fig:fbeta_avg}. \addtext{To take the average, we set the width of the moving window to $0.3\beta_0$, so that it narrows as the band limit increases, keeping the number of oscillations in the window roughly the same.  The finding of $\beta^{-2}$} agrees with the hypothesis for the {flat sky and continuum case}, for which e.g. Zaldarriaga (2001) \cite{Zaldarriaga2001a} argued that form of the radial function has to be ${}_\mm f(\beta>0)\simeq\beta^{-2}$ to ensure that the Fourier modes of Stokes $Q/U$ relate to those of $E/B$ merely by rotations (with no scale dependence). In addition, we find that ${}_{\md}f$ under a moving average shows a similar scaling of $\beta^{-2}$, as seen in \fig{fig:dfbeta_avg}.
\addtext{When the moving average is applied radial functions with progressively higher band limits, we see that a larger and larger fraction of angular domain matches the scaling of $\beta^{-2}$.}   Extrapolating this trend to the case of a very high band limit one expects this behavior to approach $\beta^{-2}$ across the domain $\beta\in[\epsilon,\pi-\epsilon]$ for $\epsilon \ll 1$. In essence the moving average can be understood as the averaging resulting from discrete sampling of a continuum field (one with an infinite band limit) where the width of the averaging window corresponds to the size of the pixel.

\addtext{In \fig{fig:fbeta_avg}, near the edges ($\beta \rightarrow 0,\pi$), the window function partly falls outside the domain, and this in part causes the the scaling to deviate from the power law $\beta^{-2}$.}  Note also that ${}_{\mm}f$ and ${}_{\md}f$ need to vanish at $\beta=0$ and therefore the sliding average in the vicinity of these points cannot have the same power law behavior.
%---------------------------------------------------------------------------------------------------------
%\include{generalizedeb}
\section{Generalized polarization operators and recovering standard power spectra}\label{sec:generalized_operators}  
\addtext{In this section we relate our new-found knowledge about the radial part of the kernel to the computation of the standard CMB polarization power spectra.} With this better understanding, we can write down generalized $E/B$-like fields that depend on a different radial function, even one that we specify to have compact support.
The spin symmetry constrains the the azimuthal part of the real space kernels to be of the form $\sim e^{\pm i2 \alpha}$.  The radial parts of the standard operators are determined by the sum over spherical harmonics and varies as a function of the band limit. It is here that we may potentially choose alternate forms for the radial functions to suit certain kind of analysis.

We can systematically generalize the real space operator by introducing the following harmonic space filter function:
%
\beq
\tilde{\mathcal{G}} = {\begin{bmatrix} g_{\ell}^E & 0  \\  0 & g_{\ell}^B \end{bmatrix}} \,,
\eeq
%
where the functions $g_{\ell}^E$ and $g_{\ell}^B$ represent the harmonic representation of the modified radial functions and can in the most general case be chosen to be different for $E$ and $B$ modes. To simplify discussions, we proceed by setting $g_{\ell}^E = g_{\ell}^B= g_{\ell}$. Given this harmonic function $g_{\ell}$, we can define the real space operator $\bar{O}'$ which translates Stokes $Q/U$ to $E/B$-like scalars (and the inverse operator $\bar{O}'^{-1}$) in the following manner,
% 
\begin{subequations} \label{eq:gen_qu2eb}
\beqry
{\bar O}' &=& {{}_0\mathcal{Y}} \, \tilde T^{-1} \tilde{\mathcal{G}} {{}_2\mathcal{Y}^{\ddag}} \, \bar T \,,\\
{\bar O}'^{-1}&=& \bar{T}^{-1} {{}_2\mathcal{Y}}\, \tilde{\mathcal{G}}^{-1} \tilde T {{}_0\mathcal{Y}^{\ddag}}.
\eeqry
\end{subequations}
%
The primed notation distinguishes these generalized operators from the default operators defined in \sec{sec:qu2eb} and \sec{sec:eb2qu}. We require both the forward and inverse operators to be well defined.   This constrains the choice of $\tilde{\mathcal{G}}$ to have a valid  inverse, which is important when recovering the standard CMB power spectra. The radial parts of this generalized operator and it's inverse are given by the following expressions,
%
\begin{subequations}\label{eq:generalized_radial_kernel}
\beqry 
G_{QU \rightarrow EB}(\beta) &=& G(\beta) = \sum _{\ell=2} ^{\ell_{\rm max}} g_{\ell}\frac{2 \ell+1}{4 \pi} \sqrt{\frac{(\ell-2)!}{(\ell + 2)!}} P_{\ell}^2(\cos{\beta}) \, \label{eq:mod_rad_forward} \,, \\
G_{EB \rightarrow QU}(\beta) &=& G^{-1}(\beta) = \sum _{\ell=2} ^{\ell_{\rm max}} g_{\ell}^{-1}\frac{2 \ell+1}{4 \pi} \sqrt{\frac{(\ell-2)!}{(\ell + 2)!}} P_{\ell}^2(\cos{\beta}) \,. \label{eq:mod_rad_inverse}
\eeqry
\end{subequations}
%
The default radial function is just a special case resulting from the choice $\tilde{\mathcal{G}}=\mathbb{1}$ ($g_{\ell}=1$), in which case $\tilde{\mathcal{G}}^{-1}=\tilde{\mathcal{G}}$ and therefore $G^{-1}(\beta) = G(\beta)={{}_{\mm}f}$.
%
\begin{figure}[!t] 
\centering
\subfigure[\label{fig:gl_gbeta}]{\includegraphics[width=0.48\columnwidth]{Gbeta_for_different_gl_lmax1536.pdf}}
\subfigure[\label{fig:glbl}]{\includegraphics[width=0.48\columnwidth]{gl_for_different_gbeta_lmax1536.pdf}}
\subfigure[\label{fig:gl_bbeta}]{\includegraphics[width=0.48\columnwidth]{Bbeta_for_gl_lmax1536.pdf}}
\caption{\textit{Top left:} The green line depicts the default radial kernel $f(\beta)$ defined in \eq{eq:qu2eb_gen_kernel}, multiplied by an apodized step function $\theta(3 \beta_0)$. The blue and orange lines depict the modified radial function resulting the beam harmonics $b_{\ell}$ corresponding to Gaussian beams with fwhm=15 \& 12 arc-minutes respectively. The red curve depicts an example modified radial function: $G(\beta)=\mathcal{A} \beta^{n_o} \exp{\left[ -\left( {(\beta-\beta_s)}{/\sqrt{2} \sigma} \right)^{n_1} \right]}$ with parameters set to the following values $[n_0=1;\, \beta_s=0 ;\, \sigma = 0.004 ;\, n_1=1.5]$. \textit{Top right:} The harmonic representation of the respective radial functions as indicated by the legend. The dashed curves of the corresponding color depict the inverse of the harmonic functions. \textit{Bottom:}  The normalized beam function $b(\beta)$ evaluated from interpreting the respective harmonic functions as those corresponding to an effective beam applied to $E/B$. }
\label{fig:example_gbeta}
\end{figure}
%
